---
title: 'Class Activity #2'
author: "Write Your Name"
date: '2021-01-27'
output:
  pdf_document: default
  html_document: default
---
Section 2.3: Linear Regression and Hypothesis Test for the Game data

1) Use the software instructions and the Game data to create indicator variables where $x=1$ represents the color distracter game and $x=0$ represents the standard game. Develop a regression model using Time as the response and the indicator variable as the explanatory variable.

```{r}
Game <- read.csv("C2 Games1.csv")
```
Since Standard and Color are not numbers, we turn the Type column into a boolean array. We then multiply that array by one in order to turn the trues into 1 and false into 0.
```{r}
D <- (Game$Type == "Color")*1 # boolean array for color = 1, standard = 0
```
Now that we have numerical data to work with, we can fit a linear model (lm) to our data and calling the summary function on our linear model Fit, we see very useful information about Residuals and coefficients. 
```{r}
Fit <- lm(Game$Time~D) # least square linear regression 

summary(Fit)

```

2) Use statistical software to calculate the t-statistic and p-value for the hypothesis tests. $H_o:\beta_1=0$ versus $H_a: \beta_1\neq 0$. In addition, construct a 95% confidence interval for $\beta_1$. Based on these statistics, can you conclude that the coefficient, $\beta_1$, is significantly different from zero? 

From the summary above, we find that the Pr(>|t|) for $\beta_1$ is less than our $\alpha = .05$, thus, we must reject the null hypothesis. In order to  construct a $95\%$ confidence interval, we must take the estimate for $\beta_1$ and add/subtract the $t_{df, \frac{\alpha}{2}}sd(\hat\beta)$ which can also obtain from the summary above. 
```{r}
Lower <- 2.55 + qt(.025, df = 38) * 1.1154 

Upper <- 2.55 - qt(.025, df = 38) * 1.1154

cat(sprintf("Lower: %.8f\n", Lower))
cat(sprintf("Upper: %.8f", Upper))
```

3) Repeat the two previous questions, but use an indicator where $x=1$ represents the standard game and $x=0$  represents the color distracter game. Compare the regression line, hypothesis test, and p-value to those from the previous questions. 

Repeating the steps from above:
```{r}
D1 <- (Game$Type == "Standard")*1

Fit1 <- lm(Game$Time~D1) # least square linear regression 

summary(Fit1)

Lower <- -2.55 + qt(.025, df = 38) * 1.1154 

Upper <- -2.55 - qt(.025, df = 38) * 1.1154 

cat(sprintf("Lower: %.8f\n", Lower))
cat(sprintf("Upper: %.8f", Upper))

```
Here we see that everything is the same except that now our interval is negative. 

4) Calculate the residuals from the regression line in question 1. Plot a histogram of the residual (or create a normal probability plot of the residuals). In addition, create a residual versus order plot and use the informal test to determine if the equal variance assumption is appropriate for this study. Compare these plots to the residual plots created for the two sample t-test.

Earlier, we fit a linear model onto our data using the lm() function and called it Fit. Looking at the summary of Fit, we find that Fit has many variables which are very useful to us, one of which is the residuals. Therefore, since Fit contains information about our residuals we can simply call it from our Fit variable. 
```{r}
#using variables from questions 1 and 2 not from 3
Resid <- Fit$residuals
Resid
```
Histogram of the residuals:
```{r}
hist(Resid)
qqnorm(Resid)
qqline(Resid)

```
A plot of the residuals according to the studentID, to make sure that there is no pattern which exists in our residuals. 
```{r}
ID <- Game$studentID

plot(Resid~ID)
points(Resid~ID, type = "l") #you can use any order 
mean(Resid)
```

Using the informal test to check variance equivalence, we find that since our max var over our min var is less than 4, we assume our variance in the two groups are equal: 
```{r}
Standard <- subset(Game, Game$Type == "Standard")
Color <- subset(Game, Game$Type == "Color")
max(var(Standard$Time), var(Color$Time)) / min(var(Standard$Time), var(Color$Time)) < 4

```

5) Create a scatterplot with the regression line in question 1. Use the graph to give an interpretation of the slope and y-intercept, $b_1$ and $b_0$ in the context of the game study.

```{r}
plot(Game$Time~D, ylab="Time", xlab="Type", main="Regression for Comparing Two Populations")
abline(Fit, col = "Blue")
```
$\beta_1$ is the rate of change of $y$ for every $D_i$ = 1 and $\beta_0$ is the average of y when $D_i = 0$ 
